### 2022-08-05 supp/01-size.class.Rmd
<init analysis>

# Define size classes

## Load packages + data

```{r load.packages.szclass, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# load required analysis packages
library(here)
library(StereoMorph)
library(geomorph)
library(tidyverse)
library(wesanderson)
library(ggplot2)

# read data and define number of sLMs ----
shapes <- readShapes("shapes")
shapesGM <- readland.shapes(shapes, nCurvePts = c(15, 10, 15))

# read qualitative data
qdata <- read.csv("qdata.csv", 
                  header = TRUE, 
                  row.names = 1)
```

## Generalised Procrustes Analysis

Landmark data were aligned to a global coordinate system [@RN11622;@RN11623;@RN11563], achieved through generalized Procrustes superimposition [@RN478] performed in R 4.1.3 [@R] using the `geomorph` library v. 4.0.3 [@RN11530;@RN1774;@RN9565]. Procrustes superimposition translates, scales, and rotates the coordinate data to allow for comparisons among objects [@RN11564;@RN478]. The `geomorph` package uses a partial Procrustes superimposition that projects the aligned specimens into tangent space subsequent to alignment in preparation for the use of multivariate methods that assume linear space [@RN1646;@RN11563].

```{r outlines.szclass, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# gpa ----
Y.gpa <- gpagen(shapesGM, print.progress = FALSE)

# plot
plot(Y.gpa)

# geomorph data frame ----
gdf <- geomorph.data.frame(shape = Y.gpa$coords, 
                           size = Y.gpa$Csize,
                           site = qdata$site)

# add centroid size to qdata
qdata$csz <- Y.gpa$Csize
```

## Assign size class

```{r setwd.szclass ld define variables 3, echo=TRUE, eval = FALSE}
#select vars
pre.data <- qdata %>% 
  select(site:csz)

# millsap subset----
millsap <- filter(pre.data,
                  site == "Millsap Cache")

# calculate mean centroid size
mean.csz <- mean(millsap$csz)

# assign L/S varieties based on mean centroid size
lcheck2 <- millsap$csz >= mean.csz
lcheck3 <- millsap$csz >= 0 & millsap$csz < mean.csz

millsap$size <- ""
millsap$size <- ifelse(lcheck2, "L", millsap$size)
millsap$size <- ifelse(lcheck3, "S", millsap$size)

# an13 subset----
an13 <- subset(pre.data,
               site == "Jowell Farm")

# calculate the mean length
mean.csz <- mean(an13$csz)

# assign L/S varieties based on mean length
lcheck2 <- an13$csz >= mean.csz
lcheck3 <- an13$csz >= 0 & an13$csz < mean.csz

an13$size <- ""
an13$size <- ifelse(lcheck2, "L", an13$size)
an13$size <- ifelse(lcheck3, "S", an13$size)

# aggregate data
mx <- bind_rows(millsap, an13)

# join site and size class
qdata2 <- mx %>% 
  unite("merged", site, size, remove = FALSE)

# output csv for use in gm analysis
write.csv(qdata2, "qdata.merged.csv")
```