### 2022-10-31 ms/ms/ms.Rmd
<init ms>

---
title: |
  Size matters: Divergent morphology in elliptical bifaces from sites in the American Southeast articulates with distinct local reduction practices
type: ARTICLE TEMPLATE
author:
  - name: Robert Z. Selden, Jr.
    affil: a
    email: zselden@sfasu.edu
  - name: John E. Dockall
    affil: b
    email: john.dockall@stantec.com
  - name: David H. Dye
    affil: c
    email: daviddye@memphis.edu
affiliation:
  - num: a
    address: |
      Heritage Research Center, Stephen F. Austin State University; Department of Biology, Stephen F. Austin State University; Texas Archeological Research Laboratory, The University of Texas at Austin; and Cultural Heritage Department, Jean Monnet University 
  - num: b
    address: |
      Stantec, Inc.
  - num: c
    address: |
      Department of Earth Sciences, The University of Memphis
bibliography: interactcadsample.bib
appendix: appendix.tex
abstract: |
  Elliptical bifaces are prevalent at Mississippian sites throughout the American Southeast. Those from Millsap Cache and Jowell Farm (41AN13) comprise two of the largest samples of this ill-understood stone tool from the ancestral Caddo area. Elliptical bifaces from Millsap Cache were produced using Kay County flint, while those from Jowell Farm were manufactured using Edwards chert, providing for an empirical test of morphological differences as a function of raw material. The sample of elliptical bifaces was divided into two size classes; one conceptually reflective of production (large), and the other with local reduction practices (small). Size classes were used to assess whether modifications by Caddo knappers may have yielded similar—convergent—biface shape in the small class. Size classes were also used to test the hypothesis that greater morphological variation would be apparent in the small class due to idiosyncratic responses related to local retouch practices. Results demonstrate that elliptical biface shape does not differ by raw material, but size does. This suggests that a shared, and consistent, mental template was maintained independent of biface size. The subsequent analysis of elliptical biface morphology by size class demonstrated that size does not differ by raw material in the large class, but in the small class, it does. This finding supports the argument that elliptical biface morphology diverges through local reduction practices. As expected, greater shape diversity occurs in the small class, where Jowell Farm bifaces were found to be more morphologically diverse than those from Millsap Cache. Distinct local reduction practices are advanced as the driver of extant morphological differences found in elliptical bifaces.
keywords: |
  American Southeast; Caddo; NAGPRA; archaeology; ovoid biface; Jowell knife; Jowell knives; lithics; oblique parallel flaking; archaeoinformatics; museum studies; digital humanities; non-western art history; geometric morphometrics; STEM; STEAM
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: rticles::tf_article
---

\begin{quote}
\textit{This process of comparison, of recognising in one form a definite permutation or deformation of another, apart altogether from a precise and adequate understanding of the original 'type' or standard of comparison, lies within the immediate province of mathematics, and finds its solution in the elementary use of a certain method of the mathematician} \citep{RN7522}.
\end{quote}

# Introduction

\begin{figure}\centering
\includegraphics[width=\linewidth]{figs/map.png}
\caption{Location of Millsap Cache, Jowell Farm, and other sites mentioned in the manuscript. Elliptical bifaces have been found in every major river basin in the ancestral Caddo area (white).}
\label{fig:map}
\end{figure}

\begin{figure}\centering
\includegraphics[width=\linewidth]{figs/elliptical.illustration.png}
\caption{Illustrated elliptical bifaces from a, an unknown location; b, Jowell Farm; and c, Belcher Mound highlighting oblique parallel flaking used in production. The specimen from Belcher Mound is a potential preform exhibiting oblique parallel removals consistent with elliptical bifaces from other sites in the region.}
\label{fig:illustrated}
\end{figure}

# Methods

To assess the variable impacts that Caddo retouch practices may have had on general morphology, the sample of elliptical bifaces was binned into two size classes using centroid size. Due to potential morphological differences associated with the use of two distinct raw materials (Kay County flint and Edwards chert), the sample was subset by site prior to calculating the mean centroid size for each. Bifaces larger than the mean were then assigned to the large category, those smaller than the mean were assigned to the small category, and the two datasets were then joined in advance of analysis ([Supplementary Materials, Ch. 1](https://seldenlab.github.io/elliptical.bifaces/define-size-classes.html)).

Elliptical bifaces from the Millsap Cache and Jowell Farm were used to examine whether biface morphology remains stable or expresses morphological variability by site/raw material ([Supplementary Materials, Ch. 2](https://seldenlab.github.io/elliptical.bifaces/gm---siteraw-material.html)). Prior to landmarking, elliptical bifaces were oriented with the most heavily retouched edge at top right (Figure  \ref{fig:elliptical}). Some bifaces include multiple areas of heavy retouch at the top and bottom of the same lateral edge (Figure \ref{fig:elliptical}:b, c, d, and e), while others (Figure \ref{fig:elliptical}:f) include one retouched edge at top right and another at bottom left.

\begin{figure}\centering
\includegraphics[width=\linewidth]{figs/ellipticalbifaces.png}
\caption{Selected elliptical bifaces from the Jowell Farm site highlighting the most heavily retouched area identified using a modified version of the flaking index developed and advanced by \citet{RN11099}, \citet{RN9242}, and \citet{RN11098}.}
\label{fig:elliptical}
\end{figure}

To identify which edge was most heavily retouched, we employed a modified approach to the flaking index initially developed by \citet{RN11099}, and later advanced and refined by \citet{RN9242}, then \citet{RN11098}. The modified approach uses counts of flake scars from each edge in the two most heavily worked areas, paired with a measure of edge length inclusive of curvature, which was calculated using ImageJ (Abramoff et al. 2004, Rasband 1997, Schneider et al. 2012). The number of flake scars was then divided by the length of the worked edge, with the most heavily worked edge identified by the greater value.

## Geometric morphometrics

The same landmark/semilandmark configuration was used for both analyses, and the landmarking protocol employs three landmarks; two horizontal tangents (top/bottom), and the third placed at the furthest extent of the retouched edge bearing the heaviest amount of retouch (see Figure \ref{fig:elliptical}). Equidistant semilandmarks were placed between each landmark, and all landmarks and semilandmarks were applied using R 4.2.1 (R Core Development Team, 2022) and the StereoMorph package (Olsen and Westneat 2015).

Landmark data were aligned to a global coordinate system (Bookstein et al. 1999, Gunz et al. 2005, Kendall 1981, 1984, Slice 2001), achieved through generalised Procrustes superimposition (Bookstein 1986, Rohlf and Slice 1990, Rohlf 1999) in R using the geomorph (Adams et al. 2017, Adams and Otarola-Castillo 2013, Baken et al. 2021) and RRPP packages (Adams and Collyer 2015, Collyer and Adams 2018). Procrustes superimposition translates and rotates the coordinate data to allow for comparisons among objects, while also scaling each biface using unit-centroid size—the square root of the sum of squared distances from each landmark to the specimen’s centroid (Chapman 1990, Dryden and Mardia 1998, Gower 1975, Rohlf and Slice 1990). The geomorph package uses a partial Procrustes superimposition that projects the aligned specimens into tangent space subsequent to alignment in preparation for the use of multivariate methods that assume linear space (Dryden and Mardia 1993, Kent and Mardia 2001, Rohlf 1999, Slice 2001).

# Acknowledgement(s) {-}

Our thanks to the Caddo Nation of Oklahoma, the Caddo Nation Tribal Council, Tribal Chairman, and Tribal Historic Preservation Office for permission and access to NAGPRA and previously repatriated collections. Our gratitude is also extended to Marybeth Tomka and Lauren Bussiere at the Texas Archeological Research Laboratory for their assistance with access to the bifaces and associated records, to Sergio Ayala and the Gault School of Archaeological Research for access to the UV light, and to Scott Hammerstedt and Debra K. Green at the Oklahoma Archeological Survey for their assistance with records requests. Thanks also to John Harman for access to the DStretch plugin for ImageJ that was useful in the analysis of flake scars, and to Harry J. Shafer, Hiram F. (Pete) Gregory, Christian S. Hoggard, and David K. Thulman for their comments and constructive criticisms on the ongoing analyses of Caddo biface morphology, as well as Emma Sherratt, Kersten Bergstrom, Lauren Butaric, Dean C. Adams, and Michael L. Collyer for their constructive criticisms, general comments, and suggestions throughout the development of this research program.

# Funding {-}

Components of this analytical work flow were developed and funded by a Preservation Technology and Training grant (P14AP00138) to RZS from the National Center for Preservation Technology and Training (NCPTT), and additional grants to RZS from the Caddo Nation of Oklahoma, National Forests and Grasslands in Texas (15-PA-11081300-033) and the United States Forest Service (20-PA-11081300-074). Funding to analyse the bifaces from Millsap Cache and Jowell Farm was made possible by support from the Heritage Research Center at Stephen F. Austin State University. A Research Support Fund Grant from the Texas Archeological Society provided the tool needed to illustrate the bifaces.

# Data management

Reproducibility—the ability to recompute results—and replicability—the chances other experimenters will achieve a consistent result—are two foundational characteristics of successful scientific research (Leek and Peng 2015). The analysis code associated with this project can be accessed through the [Supplementary Materials](https://seldenlab.github.io/elliptical.bifaces/), is available through the GitHub repository [https://github.com/seldenlab/elliptical.bifaces](https://github.com/seldenlab/elliptical.bifaces), and digitally curated on the Open Science Framework [DOI 10.17605/OSF.IO/PH25W](https://osf.io/ph25w/). The reproducible nature of this enterprise provides a means for others to critically assess and evaluate the various analytical components (Gray and Marwick 2019; Peng 2011; Gandrud 2014), which is a necessary requirement for the production of reliable knowledge.

Reproducibility projects in psychology and cancer biology are impacting current research practices across all domains. Examples of reproducible research are becoming more abundant in archaeology (Marwick 2016; Ivanovaite et al. 2020; Selden Jr., Dockall, and Shafer 2018; Selden Jr., Dockall, and Dubied 2020; Selden Jr et al. 2021a; Selden 2022), and the next generation of archaeologists are learning those tools and methods needed to reproduce and/or replicate research results (Marwick et al. 2019). Reproducible and replicable research work flows are often employed at the highest levels of humanities-based inquiries to mitigate concern or doubt regarding proper execution, and is of particular import should the results have—explicitly or implicitly—a major impact on scientific progress (Peels and Bouter 2018).