### 2022-12-21 supp/dance.qmd
<init lit>

---
title: "Literature: Dance"
bibliography: [book.bib]
---

The dataset used in this analysis was harvested from the [Web of Science](https://clarivate.com/webofsciencegroup/solutions/web-of-science/), includes all 611 publications returned for the query **_dance AND (archaeology OR archeology OR anthropology)_**, and uses the `bibliometrix` package for analysis.

```{r bibliometrix, echo=TRUE, warning=FALSE}
# install
#devtools::install_github("massimoaria/bibliometrix")

# load
library(here)
library(bibliometrix)
library(reshape2)
library(ggplot2)

# build data frame
df <- convert2df(file = "savedrecs.bib",
                 dbsource = "wos",
                 format = "bibtex")
```

## Descriptive analysis

```{r summary, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
results <- biblioAnalysis(df, 
                          sep = ";")
options(width = 100)
s <- summary(object = results,
             k = 10,
             pause = FALSE)

plot(x = results,
     k = 10,
     pause = FALSE)
```

### Attributes of the local network

```{r local.attr, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# calculate citations in local network
CR <- localCitations(df, sep = ";")

# top 10 cited authors in local network
knitr::kable((CR$Authors[1:10,]), "simple")

# top 10 cited papers in local network
knitr::kable((CR$Papers[1:10,]), "simple")
```

## Most cited

### Most cited articles

```{r most.cited.refs, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# most cited references in global network
mcr <- citations(df,
                 field = "article",
                 sep = ";")

cbind(mcr$Cited[1:10])
```

### Most cited authors

```{r most.cited.auths, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# most cited authors in global network
mcr <- citations(df,
                 field = "author",
                 sep = ";")

cbind(mcr$Cited[1:10])
```

### Author dominance ranking

```{r auth.dom, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
dom <- biblioAnalysis(df)
dom.r <- dominance(dom)

knitr::kable(dom.r, "simple")
```

### Productivity over time

```{r prod, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
topAU <- authorProdOverTime(df, k = 10, graph = TRUE)
```

## Historiograph direct citation linkages

The historiographic map is a chronological network map of the most relevant direct citations resulting from this bibliographic collection.

```{r hdcn, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# historical citation network
options(width = 100)
histResults <- histNetwork(df,
                           min.citations = 20,
                           sep = ";")

# plot historical co-citation network
hnet <- histPlot(histResults,
                 n = 35,
                 size = 4,
                 labelsize = 1)
```

### ID and DE keyword associations

```{r key.assocx, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
list <- keywordAssoc(df,
                     sep = ";",
                     n = 10)

list[[1]][1:10]
```

### Yearly occurrences of top keywords/terms

#### Authors' keywords

```{r key.growth, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
topKW = KeywordGrowth(df,
                      Tag = "DE",
                      sep = ";",
                      top = 10,
                      cdf = TRUE)

topKW

# plot results
key.plot = melt(topKW,
                id ='Year')

ggplot(key.plot, aes(Year,
                     value,
                     group = variable,
                     color = variable)) + 
  geom_line()
```

#### Publisher's keywords

```{r key.growth2x, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
topKW = KeywordGrowth(df,
                      Tag = "ID",
                      sep = ";",
                      top = 10,
                      cdf = TRUE)

topKW

# plot results
key.plot = melt(topKW,
                id ='Year')

ggplot(key.plot, aes(Year,
                     value,
                     group = variable,
                     color = variable)) + 
  geom_line()
```

## Conceptual structure

### Co-word analysis

The co-word analysis maps the conceptual structure of a research domain using the co-occurrence of author keywords in the bibliographic collection.

#### Author's keywords

```{r co.word, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# using authors keywords
cw <- conceptualStructure(df,
                          field = "DE",
                          method = "MCA",
                          minDegree = 5,
                          clust = "auto",
                          stemming = FALSE,
                          labelsize = 8)
```

#### Publisher's keywords

```{r co.word2i, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# using publishers keywords
cw <- conceptualStructure(df,
                          field = "ID",
                          method = "MCA",
                          minDegree = 5,
                          clust = "auto",
                          stemming = FALSE,
                          labelsize = 8)
```

## Thematic mapping

From [@RN20996, 150-151]:

* Themes in the upper-right quadrant are both well developed and important for the structuring ofa research field. They are known as the motor-themes of the specialty, given that they present strong centrality and high density. The placement of themes in this quadrant implies that they are related externally to concepts applicable to other themes that are conceptually closely related.
* Themes in the upper-left quadrant have well developed internal ties but unimportant external ties and so are of only marginal importance for the field. These themes are very specialized and peripheral in character.
* Themes in the lower-left quadrant are both weakly developed and marginal. The themes of this quadrant have low density and low centrality, mainly representing either emerging or disappearing themes.
* Themes in the lower-right quadrant are important for a research field but are not developed. So, this quadrant groups transversal and general, basic themes.

### Authors' keywords

```{r thematic.map, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# keyword map
map1 = thematicMap(df,
                   field = "DE",
                   n = 1000,
                   minfreq = 4,
                   ngrams = 1,
                   stemming = FALSE,
                   size = 0.3,
                   n.labels = 1,
                   community.repulsion = 0.1,
                   cluster = "walktrap",
                   repel = TRUE)
# plot map
plot(map1$map)
```

### Publisher's keywords

```{r thematic.map2b, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# keyword map
map2 = thematicMap(df,
                   field = "ID",
                   n = 1000,
                   minfreq = 4,
                   ngrams = 1,
                   stemming = FALSE,
                   size = 0.6,
                   n.labels = 1,
                   community.repulsion = 0.1,
                   cluster = "walktrap",
                   repel = TRUE)
# plot map
plot(map2$map)
```

## Social structure

### Author collaboration

Scientific collaborations are plotted where nodes are authors and links are co-authorships, illustrating collaborations between authors.

```{r auth.collab, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# author collaboration network
auth.collab <- biblioNetwork(df,
                             analysis = "collaboration",
                             network = "authors",
                             sep = ";")

# network plot
auth.collabnet = networkPlot(auth.collab,
                             n = 75,
                             Title = "Author Collaboration",
                             type = "mds",
                             halo = TRUE,
                             weighted = TRUE,
                             size = 5,
                             edgesize = 1,
                             labelsize = 0.5,
                             remove.isolates = TRUE)
```

### Edu collaboration

Scientific collaborations are plotted where nodes are institutions and links are co-authorships, illustrating collaborations between institutions.

```{r edu.network, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# author collaboration network
edu.collab <- biblioNetwork(df,
                            analysis = "collaboration",
                            network = "universities",
                            sep = ";")

# network plot
edu.collabnet = networkPlot(edu.collab,
                            n = 75,
                            Title = "Edu Collaboration", 
                            type = "mds",
                            halo = TRUE,
                            size = 15, 
                            size.cex = T, 
                            edgesize = 1, 
                            labelsize = 0.5, 
                            remove.isolates = TRUE)

# end of code
```